---
title: "Raw Data"
author: "Marissa Douglas"
date: "2023-07-12"
output: pdf_document
---

Load Packages
```{r}
library(tidyverse)
library(ggplot2)
library(ggimage)
library(ggpp)
library(glmmTMB)
library(sjPlot)
theme_set(theme_classic())
```

Load the CSV

```{r}
ants_wide <- read_csv("Competition experiment data - Tidy Census.csv")
```
Let's add a column of the queens as a binary true/false
```{r}
ants_wide <- ants_wide %>% mutate(queentf = ifelse(queen>0,TRUE, FALSE), broodtf = ifelse(brood == "y", TRUE, FALSE))
```

A long data set may work best for a bar plot (geom_bar)
```{r}
# create a data set by making data long
ants <- ants_wide %>% 
  pivot_longer(c(`workers`, `soldiers`, `queen`, `winged queens`, `males`), names_to = "caste", values_to = "count")
```



VISUALIZATION:


Make a graph for each colony (4 graphs). Group them by type (control/treatment). Make a histogram of the count (workers + soldiers) of ants in each nest. For each bar, the aesthetic should display worker count on the bottom and soldier count on the top. Add a symbol for the presence of queens or brood.

Let's make 399's tibble
```{r}
#make a tibble of the colony's nests
pw399 <- ants %>% 
  filter(`colony name` == "PW399", nest != "dead", nest != "outside")
```


399 graph:
source: https://stackoverflow.com/questions/12715635/ggplot2-bar-plot-with-both-stack-and-dodge
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw399 %>% filter(caste %in% c("workers", "soldiers"), nest != "base") %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill =type)) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

ggsave("colony399.png", width = 7, height = 4)
```

Let's make 437's tibble
```{r}
#make a tibble of the colony's nests
pw437 <- ants %>% 
  filter(`colony name` == "PW437", nest != "dead", nest != "outside")
```

437 graph:
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw437 %>% filter(caste %in% c("workers", "soldiers")) %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill =forcats::fct_rev(type))) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

ggsave("colony437.png", width = 7, height = 4)
```

Let's make 398's tibble
```{r}
#make a tibble of the colony's nests
pw398 <- ants %>% 
  filter(`colony name` == "PW398 (rascal)", nest != "dead", nest != "outside")
```

398 graph:
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw398 %>% filter(caste %in% c("workers", "soldiers"), nest != "base") %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill = type)) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

ggsave("colony398.png", width = 7, height = 4)
```


Let's make 400's tibble
```{r}
#make a tibble of the colony's nests
pw400 <- ants %>% 
  filter(`colony name` == "PW400 (shawshank)", nest != "dead", nest != "outside")
```

437 graph:
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw400 %>% filter(caste %in% c("workers", "soldiers")) %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill =forcats::fct_rev(type))) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

ggsave("colony400.png", width = 7, height = 4)
```




STATISTICAL ANALYSIS


Next, let's create a model to look for a preference for specific nests per treatment.

First, make a data frame that includes the sum of the workers and the soldiers
```{r}
ants_sum <- ants_wide %>% 
  filter(type != "test", nest != "dead", nest != "outside", nest != "base") %>%
  mutate(
    workerSum = workers + soldiers)
colnames(ants_sum)[1] ="colony"

```


```{r}
model.workers.bynest <- glmmTMB(workerSum ~ nest*type + 
                          (1|colony) -1,
                        family="nbinom2", data = ants_sum)

summary(model.workers.bynest)
sjPlot::tab_model(model.workers.bynest)

```
What if we were to lump rows into the regions A, BCD, E, FGH?
Add a column that assigns regions

```{r}
ants_region_assignment <- ants_sum %>% 
  mutate(region = ifelse(nest == "A", "A",
                         ifelse(nest == "B", "BCD",
                                ifelse(nest == "C", "BCD",
                                       ifelse(nest == "D", "BCD",
                                              ifelse(nest == "E", "E",
                                                     ifelse(nest == "F", "FGH",
                                                            ifelse(nest == "G", "FGH",
                                                                   ifelse(nest == "H", "FGH","none")))))))))


```

Now, combine rows of the same region. Source: https://www.r-bloggers.com/2022/08/combine-rows-with-same-column-values-in-r/
```{r}
#not what I want to do and doesn't work
ants_by_regions <- ants_region_assignment %>%
group_by(colony, type, region) %>% 
  summarize(across(c(workers, soldiers, workerSum), sum))
  
```

```{r}
model.workers.byregion <- glmmTMB(workerSum ~ region*type + 
                          (1|colony) -1,
                        family="nbinom2", data = ants_by_regions)

summary(model.workers.byregion)
sjPlot::tab_model(model.workers.byregion)

```
















Can ignore:

Practice runs:

Let's graph 399. I think I need to make vectors to use a histogram. A bar plot may offer better functionality
```{r}
#bar plot attempt
#https://statisticsglobe.com/draw-grouped-barplot-in-r
#Grouped barplot using ggplot2
#Should compare the two dates in pairs, once that data is collected. I do not know how to make the bars show soldier vs worker. I do not know how to make the key specify which is control and which is competitor. 

pw399 %>% ggplot(                                      
       aes(x = nest,
           y = count,
           fill = type)) +
  geom_bar(stat = "identity",
           position = "dodge")
```

Stacking attempt

```{r}
#my attempt:
#TODO pair and add queen and brood symbol
#vector for scale fill manual
fills <- c("workers" = "darkgoldenrod1", soldiers = "mediumpurple")
# Stacked
pw399 %>% filter(caste %in% c("workers", "soldiers")) %>% 
  ggplot(aes(fill=caste, y=count, x=nest)) + 
    geom_bar(position="stack", stat="identity")+
  scale_fill_manual(values = fills)
```