---
title: "Raw Data"
author: "Marissa Douglas"
date: "2023-07-12"
output: pdf_document
---

Load Packages
```{r}
library(tidyverse)
library(broom)
library(ggplot2)
library(ggimage)
library(ggpp)
library(glmmTMB)
library(sjPlot)
theme_set(theme_classic())
```

Load the CSV

```{r}
ants_wide <- read_csv("Competition experiment data - Tidy Census.csv")
```
Let's add a column of the queens as a binary true/false
```{r}
ants_wide <- ants_wide %>% mutate(queentf = ifelse(queen>0,TRUE, FALSE), broodtf = ifelse(brood == "y", TRUE, FALSE))
```

A long data set may work best for a bar plot (geom_bar)
```{r}
# create a data set by making data long
ants <- ants_wide %>% 
  pivot_longer(c(`workers`, `soldiers`, `queen`, `winged queens`, `males`), names_to = "caste", values_to = "count")
```



VISUALIZATION:


Make a graph for each colony (4 graphs). Group them by type (control/treatment). Make a histogram of the count (workers + soldiers) of ants in each nest. For each bar, the aesthetic should display worker count on the bottom and soldier count on the top. Add a symbol for the presence of queens or brood.

Let's make 399's tibble
```{r}
#make a tibble of the colony's nests
pw399 <- ants %>% 
  filter(`colony name` == "PW399", nest != "dead", nest != "outside")
```


399 graph:
source: https://stackoverflow.com/questions/12715635/ggplot2-bar-plot-with-both-stack-and-dodge
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw399 %>% filter(caste %in% c("workers", "soldiers"), nest != "base") %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill =type)) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

#ggsave("colony399.png", width = 7, height = 4)
```

Let's make 437's tibble
```{r}
#make a tibble of the colony's nests
pw437 <- ants %>% 
  filter(`colony name` == "PW437", nest != "dead", nest != "outside")
```

437 graph:
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw437 %>% filter(caste %in% c("workers", "soldiers")) %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill =forcats::fct_rev(type))) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

#ggsave("colony437.png", width = 7, height = 4)
```

Let's make 398's tibble
```{r}
#make a tibble of the colony's nests
pw398 <- ants %>% 
  filter(`colony name` == "PW398 (rascal)", nest != "dead", nest != "outside")
```

398 graph:
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw398 %>% filter(caste %in% c("workers", "soldiers"), nest != "base") %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill = type)) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = type), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

#ggsave("colony398.png", width = 7, height = 4)
```


Let's make 400's tibble
```{r}
#make a tibble of the colony's nests
pw400 <- ants %>% 
  filter(`colony name` == "PW400 (shawshank)", nest != "dead", nest != "outside")
```

437 graph:
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

pw400 %>% filter(caste %in% c("workers", "soldiers")) %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill =forcats::fct_rev(type))) + 
  geom_col(data = . %>% filter( caste=="workers"), position = position_dodge(width = 0.9), alpha = 1) +
  geom_col(data = . %>% filter( caste=="soldiers"), position = position_dodge(width = 0.9), alpha = 0.4) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 14, width = 0.9), size=8)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = forcats::fct_rev(type)), data = . %>% filter(caste %in% c("soldiers")), position = position_dodgenudge(y = 6, width = 0.9), size=6)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 120)+
  xlab("Nests")+
  ylab("Total Number of Ants")

#ggsave("colony400.png", width = 7, height = 4)
```








NEW QUESTION:Can we represent the number of ants in a nest as a proportion of the colony?

First add in population numbers

```{r}
population <- read_csv("Competition experiment data - Population at Start Summary.csv")
```
merge this with ants_wide

```{r}
ants_wide <- ants_wide %>% 
  mutate(left_join(ants_wide, population))


```

Actually only want population that was in nests.

```{r}
#nest pop for each colony

nest_population <- ants_wide %>% 
  filter(type != "test", nest != "dead", nest != "outside", nest != "base") %>%
  group_by(`colony name`, type) %>% 
  summarize(tot_nest_pop = sum(workers + soldiers))

ants_wide <- ants_wide %>% 
  mutate(left_join(ants_wide, nest_population))


ants_wide <- ants_wide %>% 
  mutate(ants_nest_prop = (workers+soldiers)/tot_nest_pop)
```

proportion based only on nest movement:


```{r}
#add difference in proportion (control-treatment) per nest
#may be easiest to do in sheets then add in here

diff_in_prop <- read_csv("Competition experiment data - Diff_in_prop.csv")

ants_wide <- ants_wide %>% 
  mutate(right_join(ants_wide, diff_in_prop))


```


Graph with gradient of proportion info:

```{r}
# diff_median <- ants_wide %>% 
#   filter(!is.na(diff_in_proportion)) %>% 
#   median(diff_in_proportion)

ants_wide %>% 
  filter(!is.na(diff_in_proportion)) %>% 
  ggplot(aes(x=nest, y= `colony name`, color = diff_in_proportion)) + geom_point(size=12) +
  scale_color_gradient2(low="mediumpurple",mid="white",high="darkgoldenrod1", midpoint =
  0, guide=guide_colorbar(direction="horizontal",title.position="top"))
```






STATISTICAL ANALYSIS


Next, let's create a model to look for a preference for specific nests per treatment.

First, make a data frame that includes the sum of the workers and the soldiers
```{r}
ants_sum <- ants_wide %>% 
  filter(type != "test", nest != "dead", nest != "outside", nest != "base") %>%
  mutate(
    workerSum = workers + soldiers)
colnames(ants_sum)[1] ="colony"

```

Without interaction: 
```{r}
model.workers.bynest.nointeracton <- glmmTMB(workerSum ~ nest+type + 
                          (1|colony) -1,
                        family="nbinom2", data = ants_sum)

summary(model.workers.bynest.nointeracton)
sjPlot::tab_model(model.workers.bynest.nointeracton)

```

with interaction:
```{r}
model.workers.bynest <- glmmTMB(workerSum ~ nest*type + 
                          (1|colony) -1,
                        family="nbinom2", data = ants_sum)

summary(model.workers.bynest)
sjPlot::tab_model(model.workers.bynest)

```


Can ANOVA give us more insight into the significance of the GLMM models (no interaction vs interaction?)?

```{r}
anova(model.workers.bynest.nointeracton,
  model.workers.bynest)
```

What if we were to lump rows into the regions A, BCD, E, FGH?
Add a column that assigns regions

```{r}
ants_region_assignment <- ants_sum %>% 
  mutate(region = ifelse(nest == "A", "A",
                         ifelse(nest == "B", "BCD",
                                ifelse(nest == "C", "BCD",
                                       ifelse(nest == "D", "BCD",
                                              ifelse(nest == "E", "E",
                                                     ifelse(nest == "F", "FGH",
                                                            ifelse(nest == "G", "FGH",
                                                                   ifelse(nest == "H", "FGH","none")))))))))


```

Now, combine rows of the same region. Source: https://www.r-bloggers.com/2022/08/combine-rows-with-same-column-values-in-r/
```{r}
#not what I want to do and doesn't work
ants_by_regions <- ants_region_assignment %>%
group_by(colony, type, region) %>% 
  summarize(across(c(workers, soldiers, workerSum), sum))
  
```

```{r}
model.workers.byregion <- glmmTMB(workerSum ~ region*type + 
                          (1|colony) -1,
                        family="nbinom2", data = ants_by_regions)

summary(model.workers.byregion)
sjPlot::tab_model(model.workers.byregion)

```
GLMM on soldiers

```{r}
model.soldiers.bynest <- glmmTMB(soldiers ~ nest*type + 
                          (1|colony) -1,
                        family="nbinom2", data = ants_sum)

summary(model.soldiers.bynest)
sjPlot::tab_model(model.soldiers.bynest)

```
GLMM on brood

```{r}
ants_brood_binary <- ants_sum %>% 
  filter(!is.na(brood)) %>% 
  mutate(brood_binary = ifelse(brood == "y", 1, 0))
```

Why are there infinities?

```{r}
model.brood.bynest <- glmmTMB(brood_binary ~ nest*type + 
                          (1|colony) -1,
                        family="nbinom2", data = ants_brood_binary)

summary(model.brood.bynest)
sjPlot::tab_model(model.brood.bynest)

```





A long data set may work best for a bar plot (geom_bar)
```{r}
# create a data set by making data long
ants <- ants_wide %>% 
  pivot_longer(c(`workers`, `soldiers`, `queen`, `winged queens`, `males`, `workers_prop`, `soldiers_prop`), names_to = "caste", values_to = "count")
```


Let's make 399's tibble
```{r}
#make a tibble of the colony's nests
pw399 <- ants %>% 
  filter(`colony name` == "PW399", nest != "dead", nest != "outside")
```


399 graph:
source: https://stackoverflow.com/questions/12715635/ggplot2-bar-plot-with-both-stack-and-dodge
```{r}
#vector for scale fill manual
fills <- c("control" = "darkgoldenrod1", "treatment" = "mediumpurple")

prop_graph_399<-pw399 %>% filter(caste %in% c("workers_prop", "soldiers_prop"), nest != "base") %>%
  group_by(type, nest) %>% 
  mutate(cum_tot = cumsum(count)) %>% 
  ggplot(aes(nest, cum_tot, fill =type)) + 
  geom_col(data = . %>% filter( caste=="workers_prop"), position = position_dodge(width = 0.1), alpha = .1) +
  geom_col(data = . %>% filter( caste=="soldiers_prop"), position = position_dodge(width = 0.1), alpha = 0.1) +
  geom_text(aes(label=ifelse(queentf,"â™›", ""), color = type), data = . %>% filter(caste %in% c("soldiers_prop")), position = position_dodgenudge(y = .015, width = 0.8), size=.2)+
  geom_text(aes(label=ifelse(broodtf,"ðŸ¼", ""), color = type), data = . %>% filter(caste %in% c("soldiers_prop")), position = position_dodgenudge(y = .001, width = 0.8), size=.1)+
  geom_tile(aes(y=NA_integer_, alpha = factor(caste))) + 
  scale_alpha_manual(values = c(0.4,1), name = "caste")+
  scale_fill_manual(values = fills)+
  scale_color_manual(values = fills)+
  ylim(0, 0.2)+
  xlab("Nests")+
  ylab("Proportion of Ants")

prop_graph_399

#ggsave("colony399.png", width = 7, height = 4)
```
















Can ignore:

Practice runs:

Let's graph 399. I think I need to make vectors to use a histogram. A bar plot may offer better functionality
```{r}
#bar plot attempt
#https://statisticsglobe.com/draw-grouped-barplot-in-r
#Grouped barplot using ggplot2
#Should compare the two dates in pairs, once that data is collected. I do not know how to make the bars show soldier vs worker. I do not know how to make the key specify which is control and which is competitor. 

pw399 %>% ggplot(                                      
       aes(x = nest,
           y = count,
           fill = type)) +
  geom_bar(stat = "identity",
           position = "dodge")
```

Stacking attempt

```{r}
#my attempt:
#TODO pair and add queen and brood symbol
#vector for scale fill manual
fills <- c("workers" = "darkgoldenrod1", soldiers = "mediumpurple")
# Stacked
pw399 %>% filter(caste %in% c("workers", "soldiers")) %>% 
  ggplot(aes(fill=caste, y=count, x=nest)) + 
    geom_bar(position="stack", stat="identity")+
  scale_fill_manual(values = fills)
```